---
title: "Milestone Report"
author: "Camilla Nawaz"
date: "4/4/2017"
output: pdf_document

---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Milestone Report: Airbnb Ratings

## Background

On the travel site Airbnb, guests submit reviews of their experience staying in a home listed on the site. After their stay is over, the guest can submit a review of their experience, including several factors that add up to a summary review score. These reviews can impact the listing’s popularity and visibility and are an indicator of the customer base's overall experience using the site. However, it appears that a large majority of reviews fall above four stars, cheapening the value of highly-reviewed listings. In order to facilitate good experiences for guests, there are insights to be gained from analyzing the rating system. Hopefully, analyzing data from several cities with large numbers of Airbnb listings can not only help predict negative reviews, but also make positive reviews more meaningful

## The Data

The data was obtained from the site Inside Airbnb, a public site that provides data scraped from Airbnb’s site (I did not conduct any of the scraping). Each city’s data is contained in a separate .csv file. I have primarily conducted research on the datasets containing detailed summaries of each home listing; however, if time resources allow, I may also consider using text analytics to evaluate datasets containing the text of the reviews submitted by guests.

### Important Fields

The datasets each have roughly 92 variables (certain cities have a couple more or less). For the purpose of this project, the most important fields will be those relating to the price per night of the listing, room type, the review scores, the number of reviews, host acceptance rate, zipcode, and when the host joined. Notably, within the review scores, there are several fields, including accuracy, cleanliness, check-in, communication, location, and value.

### Limitations to the Data

The data is from a particular scraping date, and some of the datasets are older than a year. In this project, I assume that there will be no statistically significant difference between the current data and the slightly older scraped data, although ideally, analysis would involve the most current datasets. Additionally, not all cities or locations with Airbnb listings have data on the Inside Airbnb site; the site is not exhaustive.

### Cleaning & Wrangling

There was very minor wrangling that needed to be done for this data. I decided to keep each city’s data separate from one another, to determine if findings would differ between locations. Additionally, the original data had roughly 92 variables, many of which were not directly relevant to my research (such as several precise location variables, URLs, etc.). To keep the data more manageable and more simple, I plan to create new data frames that eliminate the extraneous variables, creating new datasets with these variables for each city:

```{r}
listings_SF_S <- listings_SF[c('id', 'interaction', 'house_rules', 'host_id', 'host_name', 'host_since', 'host_response_time', 'host_response_rate', 'host_acceptance_rate', 'host_is_superhost', 'host_listings_count', 'host_verifications', 'host_has_profile_pic', 'host_identity_verified', 'neighbourhood', 'zipcode', 'property_type', 'room_type', 'accommodates', 'bathrooms', 'bedrooms', 'beds', 'bed_type', 'amenities', 'square_feet', 'price', 'weekly_price', 'security_deposit', 'cleaning_fee', 'guests_included', 'extra_people', 'minimum_nights', 'maximum_nights', 'number_of_reviews', 'review_scores_rating', 'review_scores_accuracy', 'review_scores_cleanliness', 'review_scores_checkin', 'review_scores_communication', 'review_scores_location', 'review_scores_value', 'instant_bookable', 'require_guest_profile_picture', 'reviews_per_month')]
```

## Preliminary & Exploratory Analysis

Thus far, my analysis has included plotting visualizations in order to better understand the data. Thus far, I have found that there is an overall abundance of near-perfect ratings within each rating category. The data documenting lower ratings appears at first more sparse and less straightforwardly predictable. At this point, I plan to focus my analysis on lower ratings, using confidence intervals to select the average score rating cut-off point. To begin, I created new a new dataset containing all listings in which the average rating was less than or equal to 9/10.

```{r}
library(dplyr)
# calculate mean of all subratings, assign to new variable
listings_SF_S$mean_ratings <- mean('review_scores_accuracy', 'review_scores_cleanliness', 'review_scores_checkin', 'review_scores_communication', 'review_scores_location', 'review_scores_value')
lowratings_SF <- filter(listings_SF_S, mean_ratings < 90)
```

I also looked at histograms of all the relevant variables, hoping to determine what distributions could help determine where the most relevant and interesting analysis might lie.

```{r}
# view all variables by histogram
ggplot(data = melt(listings_SF_S), aes(value)) +
  geom_histogram() +
  facet_wrap(~variable, scales = 'free_x')
```

Additionally, I preliminarily explored how certain categorical values are linked to higher or lower ratings. For instance, when the host is listed as a superhost, their overall ratings are likely to be higher. 

```{r}
ggplot(listings_SF_S, aes(x = review_scores_rating, y = review_scores_accuracy)) +
  geom_point(position = 'jitter', alpha = 0.5, size = 0.6) +
  facet_grid(. ~ host_is_superhost) +
  xlim(0,100) +
  ylim(0,10) +
  theme_bw()
```

I had originally planned to research factors around the identity of the host (in particular, gender) and analyze how that factor contributed positive or negative ratings. However, I reasoned that the more robust analysis would lie in a broader approach to predicting positive or negative reviews, as the datasets contain many interesting variables.

## Conclusion

In conclusion, this project aims to solve two problems: 1) negative customer experiences, as expressed by negative reviews; 2) clustering around the highest review values with little variation, cheapening the value of high review scores. This analysis can lead to business solutions that may help achieve higher transparency, greater customer satisfaction, and better relations between guests and hosts.
